### 2024-12-10 README.Rmd
update readme to use stations

Here it is:
```{r}
library(sdmtools)
sin <- sdmtools::make_africa_mask(
  type = "vector",
  countries = "SGP"
)
plot(sin)
```

We're going to see how long it takes to walk home from Changi Airport. So we'll 
download the walking-only friction surface this time.

@@

And where is Changi Airport?
```{r}
changi_airport <- tibble::tibble(
  x = c(103.984),
  y = c(1.355)
)
changi_airport
```
Let's look at those.
```{r}
plot(friction_singapore)
plot(sin, add = TRUE)
points(changi_airport, pch = 19)
```

And calculate the travel time
```{r}
travel_time_sin <- calculate_travel_time(
  friction_surface = friction_singapore,
  points = changi_airport
)
travel_time_sin
```

Et voi*lah*!
```{r}
plot(travel_time_sin)
points(changi_airport, pch = 19)
plot(sin, add = TRUE)
```

### I want my plots to be nicer
```{r}
library(tidyterra)
library(ggplot2)

ggplot() +
  # plot the spatraster first
  geom_spatraster(
    data = travel_time_sin
  ) +
  theme_minimal() +
  scale_fill_whitebox_c(palette = "deep") +
  # overlay the vector outline
  geom_spatvector(
    data = sin,
    colour = "khaki3",
    linewidth = 1,
    fill = NA
  ) +
  # add the points from tibble
   geom_point(
     data = changi_airport,
     aes(x = x, y = y)
   ) +
  labs(x = NULL, y = NULL, fill = "Travel time\n(minutes)")

@@

>>>>

Here it is:
```{r}
#install.packages("geodata")
library(geodata)
sin <- gadm(
  country = "Singapore",
  level = 0,
  path = tempdir(),
  resolution = 2
)
plot(sin)
```

We're going to see how long it takes to walk home from an MRT or LRT station. So
we'll download the walking-only friction surface this time.

@@

And where are the stations? The the `stations` data set contains the latitude
and longitude of all LRT and MRT station exits [^1].
```{r}
head(stations)
plot(stations)
```
Let's look at those. Below we plot the friction surface raster, with the vector
boundary as a dashed grey line, and the station exits as grey points.
```{r}
library(tidyterra)
library(ggplot2)

ggplot() +
  geom_spatraster(
    data = friction_singapore
  ) +
  geom_spatvector(
    data = sin,
    fill = "transparent",
    col = "grey50",
    lty = 2
  ) +
  geom_point(
    data = stations,
    aes(
      x = x,
      y = y
    ),
    col = "grey60",
    size = 0.5
  ) +
  scale_fill_viridis_c(
    option = "A",
    na.value = "transparent",
    direction = -1
  ) +
  labs(
    fill = "Friction"
  )

@@

OK, now we want to calculate the walking travel time in minutes across the 
friction surface from the nearest station exit:
```{r}
travel_time_sin <- calculate_travel_time(
  friction_surface = friction_singapore,
  points = stations
)
travel_time_sin
```

Et voi*lah* --- a raster of walking time from the nearest station.
```{r}
contour(
  x = travel_time_sin,
  filled = TRUE,
  nlevels = 20,
  col = viridis::magma(19, direction = -1)
)
```

[^1]: Land Transport Authority. (2019). LTA MRT Station Exit (GEOJSON) (2024)
[Dataset]. data.gov.sg. Retrieved December 10, 2024 from
https://data.gov.sg/datasets/d_b39d3a0871985372d7e1637193335da5/view
